Hybrid platforms and bargaining power

### 2023-12-06 src/paper/application.tex
Start rewriting general results

This section presents some general result that only depend on the abstract model structure described in the previous section, and the assumptions therein.
It is remarkable that even within such a stylized setting and simple model, one can derive some interesting and non-trivial results.
\Cref{sec:example} will then build on these results to derive some stronger implications in a more specific model.

@@

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement says that the platform's profit is increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    $$\frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0$$.
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the size of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is also guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.
This decrease cannot be so strong, however, to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, axis lines=left, xlabel={$s$}]
                \addplot[name path=f, blue,ultra thick] {ln(1+x)};
                \addplot[name path=tildef, dashed, black] {x};
                \node[anchor=north west] at (axis cs: .6, 0.5) {$f(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \draw[name path=top] (axis cs:0,0.693) -- (axis cs:1,0.693);
                \draw[name path=right] (axis cs:1,0) -- (axis cs:1,0.693);

@@

                \node[anchor=north west] at (axis cs: .05, .65) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P$}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, axis lines=left, xlabel={$s$}]
                \addplot[name path=f, dashed, black] {ln(1+x)};
                \addplot[name path=tildef, blue,ultra thick] {x};
                \node[anchor=south east] at (axis cs: 0.6, 0.6) {$f(N_P', sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \draw[name path=top] (axis cs:0,1) -- (axis cs:1,1);
                \draw[name path=right] (axis cs:1,0) -- (axis cs:1,1);

@@

                \node[anchor=north west] at (axis cs: .05, .95) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P'$}
    \end{subfigure}
    \caption{Illustration of \cref{prop:share_of_platform} ($N_P < N_P'$)}
    \label{fig:increase_N_P_platform}
\end{figure}

@@

\begin{lemma}
    \label{lem:slope_at_eq}
    Let the conditions in \cref{ass:single_crossing} hold. Then for any $N_F^* > 0$ for which $\pi_F(N_P, N_F^*) = I_F$, the partial derivative with respect to the number of fringe firms is smaller than the investment cost:
    \begin{align*}
        \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_F = N_F^*} < I_F.
    \end{align*}
\end{lemma}
\begin{proof}[Proof of \cref{lem:slope_at_eq}]
    Let us use Leibniz' rule to obtain the first two derivatives of the fringe's profit with respect to the number of fringe firms:
    \begin{align*}
        \frac{\partial f(N_P, sN_F)}{\partial N_F} &= \frac{\partial}{\partial N_F} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, sN_F) \ds \right] \\
        &= \frac{\partial f(N_P, N_F)}{\partial N_F} - \int_0^1 s w(s) \frac{\partial f(N_P, sN_F)}{\partial N_F} \ds, \\
        \frac{\partial^2 f(N_P, sN_F)}{\partial N_F^2} &= \frac{\partial^2}{\partial N_F^2} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, sN_F) \ds \right] \\
        &= \frac{\partial^2 f(N_P, N_F)}{\partial N_F^2} - \int_0^1 s^2 w(s) \frac{\partial f(N_P, sN_F)}{\partial N_F^2} \ds.
    \end{align*}
    Therefore, \cref{ass:single_crossing} implies that at any $N_F$, $\pi_F(N_P, N_F)$ is either concave or decreasing.

    In the remainder, for ease of notation, let us denote partial derivatives as follows:
    \begin{align*}
        \partial_F \pi_F(N_P^*, N_F^*) \coloneqq \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_P = N_P*, N_F = N_F^*}.
    \end{align*}

@@

\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} &= \frac{\partial}{\partial N_P} \int_0^1 w(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0.
    \end{align*}
\end{proof}

>>>>

This section presents some general result that only depend on the abstract model structure described in the previous section, and the assumptions therein.
It turns out that even within such a general setting, one can derive some interesting and non-trivial results.
\Cref{sec:example} will then build on these results to derive stronger results in a more fully specific model.

@@

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement claims that -- in both the one-sided and two-sided bargaining cases -- the platform's profits are increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the slice of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is only guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.\footnote{
    In fact, one can show that this is the case if the cross-derivatives of $f$ are negative.
}
As this result shows, this decrease cannot be large enough to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, axis lines=left, xlabel={$s$}]
                \addplot[name path=f, blue,ultra thick] {ln(1+x)};
                \addplot[name path=tildef, dashed, black] {x};
                \node[anchor=north west] at (axis cs: .5, .4) {$N_C f(N_P, sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \draw[name path=top] (axis cs:0,0.693) -- (axis cs:1,0.693);
                \draw[name path=right] (axis cs:1,0) -- (axis cs:1,0.693);

@@

                \node[anchor=north west] at (axis cs: .05, .65) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P$}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick=\empty, axis lines=left, xlabel={$s$}]
                \addplot[name path=f, dashed, black] {ln(1+x)};
                \addplot[name path=tildef, blue,ultra thick] {x};
                \node[anchor=south east] at (axis cs: 0.6, 0.6) {$N_C f(N_P', sN_F)$};
                \path[name path=bottom] (axis cs:0,0) -- (axis cs:1,0);
                \draw[name path=top] (axis cs:0,1) -- (axis cs:1,1);
                \draw[name path=right] (axis cs:1,0) -- (axis cs:1,1);

@@

                \node[anchor=north west] at (axis cs: .05, .95) {Fringe ($F$)};
                \node[anchor=south east] at (axis cs: .95, .05) {Platform ($P$)};
            \end{axis}
        \end{tikzpicture}
        \caption{Profit shares with $N_P'$}
    \end{subfigure}
    \caption{Illustration of \cref{prop:share_of_platform} in the one-sided bargaining case. The right hand side figure shows a world with larger platform product variety ($N_P < N_P'$). Even though the platform's share of the total profits is smaller in relative terms in that case, it is still larger in absolute terms.}
    \label{fig:increase_N_P_platform}
\end{figure}

@@

\begin{lemma}
    \label{lem:slope_at_eq}
    Let the conditions in \cref{ass:single_crossing} hold. Then for any $N_F^* > 0$ for which $\pi_F(N_P, N_F^*) = I_F$, the partial derivative with respect to the number of fringe firms is smaller than the investment cost:
    \begin{align*}
        \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_F = N_F^*} < I_F.
    \end{align*}
\end{lemma}
\begin{proof}[Proof of \cref{lem:slope_at_eq}]
    \cref{ass:single_crossing} states that at any $N_F \geq 0$, $\pi_F(N_P, N_F)$ is either concave or decreasing.
    In the remainder, for ease of notation drop $N_C$ from the arguments of $\pi_F$ profit function as it is simply a multiplicative constant, and let us denote partial derivatives as follows:
    \begin{align*}
        \partial_F \pi_F(N_P^*, N_F^*) \coloneqq \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_P = N_P*, N_F = N_F^*}.
    \end{align*}

@@

\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F, N_C)}{\partial N_P} &= N_C \frac{\partial}{\partial N_P} \int_0^1 w(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0
    \end{align*}
    for any continuous $w(s)$.
    In particular, this holds for $w(s) = 1$ and $w(s) = s$, which correspond to assumptions \cref{ass:profit_sharing} and \cref{ass:profit_sharing_two_sided}, respectively.
\end{proof}
